---
title: "F1 Driver Info"
output: 
  flexdashboard::flex_dashboard:
    orientation: row
    vertical_layout: scroll
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(dplyr)
library(reshape2)
library(ggpubr)
library(plotly)
library(Hmisc)
library(gridExtra)
library(lubridate)
#setwd("~/Desktop/Side_proj/f1_driver_dashboard/data/")
### importing all the required files
quali <- read.csv("data/qualifying.csv")
res <- read.csv("data/results.csv")
nat <- read.csv("data/nationalityToCountry.csv")
drivers <- read.csv("data/drivers.csv")
circuits <- read.csv("data/circuits.csv")
sche <- read.csv("data/schedule.csv")
constructors <- read.csv("data/constructors.csv")
driv_st <- read.csv("data/driverStandings.csv")
### setting up input for gauge chart
drivers <- drivers %>% mutate(fullName = paste(givenName, familyName))
const_subset <- constructors %>% select(constructorId, color, name)
sche_subset <- sche %>% select(season, round, circuitId)
quali_subset <- quali %>% select(position, driverId, season, round)
res_subset <- res %>% select(driverId, season, round, position,FastestLap.rank, constructorId) 
cir_subset <- circuits %>% select(circuitId, circuitName, Location.country)
dri_subset <- drivers %>% select(driverId, nationality, fullName)

summary <- quali_subset %>% full_join(res_subset, by=c("season","round","driverId"))
summary <- summary %>% full_join(sche_subset, by=c("season","round"))
summary <- summary %>% rename(Quali_pos=position.x, Res_pos=position.y )
gauge_dat <- summary %>% full_join(const_subset, by="constructorId")

```

Sidebar {.sidebar}
=======================================================================

### Please select a driver

```{r}
selectInput("driver", "Driver:",names(table(res$driverId)>5)[which(table(res$driverId)>5)], selected = 'hamilton' )
```
This is a dashboard to display performance of formula 1 driver thhrough out their career. 

The data used to generate this dashboard is up to 2020 Hungary Grand Prix.

Data is obtained from [Ergast API](http://ergast.com/mrd/).

Source code of this project could be viewed [here](https://github.com/mkmwong/F1_driver_performance_dashboard)

P.S.(1). Only drivers with more than 5 races are included in the selection. 

P.S.(2). The earlist qualifying data is from 1994. We are not able to generate any relevant plot before that year.

P.S.(3). The earlist fastest lap data is from 2004. We are not able to generate any relevant plot before that year.

P.S.(4). WDC points each season has not been normalized by the point scheme of the year. 

Dashboard
=======================================================================

Row
-----------------------------------------------------------------------

### Age {.value-box}

```{r}

birth_year <- function(id) {
  start <- as.Date(drivers$dateOfBirth[which(drivers$driverId==id)])
  end <- as.Date(Sys.Date(), format='%d/%m/%y')
  value <- floor(time_length(interval(start,end),"years")) 
  return(value)
}
renderValueBox({
  valueBox(
    value = birth_year(input$driver),
    icon = "fa-birthday-cake"
  )
})
```

### Years in Formula 1 {.value-box}

```{r}
years_in_f1 <- function(id) {
  return(length(unique(res$season[which(res$driverId==id)])))
}
renderValueBox({
  valueBox(
    value = years_in_f1(input$driver),
    icon = "fa-calendar"
  )
})
```

### Number of WDC {.value-box}
```{r}
wdc <- function(id) {
WDC = list()
for(i in unique(driv_st$season)) {
  tmp = driv_st[which(driv_st$season==i),]
  sche_subset_tmp = sche_subset[which(sche_subset$season==i),]
  max_round = max(sche_subset_tmp$round)
  tmp = tmp[which(tmp$round==max_round & tmp$position == 1),]
  WDC = c(WDC, tmp$driverId)
}
  return(length(which(WDC==id)))
}
renderValueBox({
  valueBox(
    value = wdc(input$driver),
    icon = "fa-trophy"
  )
})
```

### Current team {.value-box}
```{r}
curr_team <- function(id){
  tmp <- full_join(res %>% select (driverId, constructorId, season), 
                   constructors %>% select(constructorId, name, color))
  tmp <- tmp %>% filter(season == '2020') %>% distinct()
  team <- tmp$name[which(tmp$driverId == id)]
  if(length(team) == 0) {
    team <- "N/A"
  }
  return(team)
}
renderValueBox({
  valueBox(
    value = curr_team(input$driver),
    icon = "fa-group"
  )
})
```

Row
-----------------------------------------------------------------------

### Number of Poles 
```{r}
renderGauge({
  get_pole_val <- function(id) {
  temp <- gauge_dat %>% filter(driverId == id)
  tab <- melt(table(temp$Quali_pos))
  dat <- list()
  if (identical(which(tab$Var1==1), integer(0))) {
    dat <- c(0,length(temp$Quali_pos))
  } else {
    dat <- c(tab$value[which(tab$Var1==1)], length(temp$Quali_pos))
  }
  return(dat)
  }
  dat <- get_pole_val(input$driver)
  rate <- dat[1]
  gauge(dat[1], min = 0, max = dat[2])
})
```

### Number of Race Wins
```{r}
renderGauge({
  get_race_val <- function(id) {
  temp <- gauge_dat %>% filter(driverId == id)
  tab <- melt(table(temp$Res_pos))
  dat <- list()
  if (identical(which(tab$Var1==1), integer(0))) {
    dat <- c(0,length(temp$Quali_pos))
  } else {
    dat <- c(tab$value[which(tab$Var1==1)], length(temp$Quali_pos))
  }
  return(dat)
  }
  dat <- get_race_val(input$driver)
  rate <- dat[1]
  gauge(dat[1], min = 0, max = dat[2])
})
```

### Number of Podiums
```{r}
renderGauge({
  get_podium_val <- function(id) {
  temp <- gauge_dat %>% filter(driverId == id)
  tab <- melt(table(temp$Res_pos))
  dat <- list()
  if (identical(which(tab$Var1==1 | tab$Var1==2 | tab$Var1==3 ), integer(0))) {
    dat <- c(0,length(temp$Quali_pos))
  } else {
    dat <- c(sum(tab$value[which(tab$Var1==1|tab$Var1==2|tab$Var1==3)]), length(temp$Quali_pos))
  }
  return(dat)
  }
  dat <- get_podium_val(input$driver)
  rate <- dat[1]
  gauge(dat[1], min = 0, max = dat[2])
})
```

### Race with Fastest Lap
```{r}
renderGauge({
  get_fl_val <- function(id) {
  temp <- gauge_dat %>% filter(driverId == id)
  tab <- melt(table(temp$FastestLap.rank))
  dat <- list()
  if (identical(which(tab$Var1==1 ), integer(0))) {
    dat <- c(0,length(temp$Quali_pos))
  } else {
    dat <- c(sum(tab$value[which(tab$Var1==1)]), length(temp$Quali_pos))
  }
  return(dat)
  }
  dat <- get_fl_val(input$driver)
  rate <- dat[1]
  gauge(dat[1], min = 0, max = dat[2])
})
```

Row
-----------------------------------------------------------------------

### Result Qualifying v.s. Race Differential

```{r}
renderPlotly({
summary <- summary %>% select(Quali_pos, Res_pos, driverId, season, circuitId)
summary <- summary %>% full_join(cir_subset, by="circuitId")

summary <- melt(summary, id=c("driverId","season","circuitId","circuitName", "Location.country"))

##### making plot for qualifying vs results #####
##### qualifying position not availabe for very old races
##### need some sort of a fix
temp <- summary %>% filter(driverId == input$driver)
max_y = max(temp$value)
seasons = sort(unique(temp$season),decreasing = TRUE) 
if (length(seasons) > 3) {
  recent_year = seasons[1:3]
} else {
  recent_year = seasons
}
temp1 <- temp %>% filter(season %in% recent_year)
p = ggplot(temp1, aes(x=variable, y=value, color = variable, text = paste0('Circuit: ', circuitName,", ",Location.country))) + theme_bw() +
  geom_boxplot(width = 0.25, color = 'black', outlier.color = NA) + geom_jitter(width = 0.25, size = 0.5)+
  facet_wrap(~ season) +
  theme( legend.position = "none") + ylim(1,max_y) + 
  xlab("") + ylab("Position") + scale_x_discrete(labels = c("Quali","Race")) + 
  scale_color_manual(values = c("grey","firebrick2")) 
p = ggplotly(p, tooltip = "text")
for(i in 1:floor((length(p$x$data)/3))){
  p$x$data[[i]]$marker$opacity = 0 
}
p[['x']][['layout']][['annotations']][[1]][['x']] <- -0.15
p
})
```

### Performance at different teams
```{r}
renderPlotly({
  summary <- res_subset %>% full_join(const_subset, by="constructorId")
  summary <- summary %>% full_join(sche_subset, by=c("season","round"))
  summary <- summary %>% full_join(cir_subset, by="circuitId")
  temp <- summary %>% filter(driverId == input$driver)
  max_y <- max(temp$position)
  temp <- temp %>% mutate(text = paste0("Season: ",season,"<br>Round: ",round,"<br>Circuit: ",circuitName, ", ", Location.country))
  d <- ggplot(temp, aes(x=reorder(name ,position,FUN = median), y=position, text = text)) + theme_bw() +
    geom_boxplot(width = 0.25, color = 'black', outlier.color = NA) + geom_jitter(width = 0.25,  color = temp$color, size = 0.5) +
    theme(legend.position = "none", axis.text.x = element_text(angle = 30, hjust = 1)) + ylim(1,max_y) + xlab("Constructor") + ylab("Position") 
  ggplotly(d)
  d = ggplotly(d, tooltip = "text")
  d$x$data[[1]]$marker$opacity = 0 
  d
})
```

### Race Performance Against Teammate
```{r}
renderPlotly({
  others <- res_subset %>% full_join(dri_subset, by="driverId")
  summary <- res_subset %>% full_join(others, by = c("season","round","constructorId"))
  summary <- summary %>% filter(driverId.x!= driverId.y)
  summary <- summary %>% mutate(pos_change = position.x - position.y)
  summary <- summary %>% full_join(const_subset, by="constructorId")
  summary <- summary %>% full_join(sche_subset, by=c("season","round"))
    summary <- summary %>% full_join(cir_subset, by="circuitId")
  temp <- summary %>% filter(driverId.x == input$driver)
  temp <- temp %>% mutate(text = paste0("Season: ",season,"<br>Round: ",round,"<br>Circuit: ",circuitName, ", ", Location.country))
  max_y = max(temp$position.x)
  c = ggplot(temp, aes(x=reorder(fullName ,pos_change,FUN = median), y=pos_change, text = text)) + theme_bw() +
    geom_boxplot(width = 0.25, color = 'black', outlier.color = NA) + geom_jitter(width = 0.25,color = temp$color,size = 0.5) +  
    theme( legend.position = "none", axis.text.x = element_text(angle = 30, hjust = 1)) + 
    ylim(-max_y,max_y) + xlab("Teammate") + ylab("Position relative to teammate") 
  ggplotly(c)
  c = ggplotly(c, tooltip = "text")
  c$x$data[[1]]$marker$opacity = 0 
  c
})
```

### Home Grand Prix Performance
```{r}
renderPlotly({
  summary <- full_join(res_subset, sche_subset, by=c("round","season"))
  summary <- full_join(summary, dri_subset, by ="driverId")
  summary <- full_join(summary, cir_subset, by ="circuitId")
  summary <- full_join(summary, nat, by="nationality")
  summary <- full_join(summary, constructors, by = "constructorId")
  summary <- summary %>% mutate(home = Location.country == country)
  summary <- summary %>% mutate(summary, sel = if_else(home == TRUE, "Yes","No"))
  temp <- summary %>% filter(driverId == input$driver)
  temp <- temp %>% mutate(text = paste0("Season:",season,"<br>Round:",round,"<br>Circuit:",circuitId))
  j = ggplot(temp, aes(x=sel, y= position)) + geom_boxplot(width = 0.25, outlier.color = NA) + 
    geom_jitter(width = 0.25,  color = temp$color, size = 0.5) + theme_bw() + 
    xlab("Home Grand Prix?") + ylab("Position") 
  j = ggplotly(j)
  j$x$data[[1]]$marker$opacity = 0 
  j
})
```

Row
-----------------------------------------------------------------------

### Circuit Performance
```{r}
renderPlotly({
  summary <- res_subset %>% full_join(sche_subset, by=c("season","round"))
  summary <- summary %>% full_join(const_subset, by="constructorId")
  summary <- summary %>% full_join(cir_subset, by="circuitId")
  summary <- summary %>% select(position, driverId, season, circuitId, circuitName, Location.country, color, round)
  ##### making plot for qualifying vs results #####
  temp <- summary %>% filter(driverId == input$driver)
  temp <- temp %>% mutate(text = paste0("Season: ",season,"<br>Round: ",round,"<br>Circuit: ",circuitName,", ",Location.country))
  max_y = max(temp$position)
  b = ggplot(temp, aes(x=reorder(circuitId,position,FUN = median), y=position, text = text)) + theme_bw() +
    geom_boxplot(width = 0.25, color = 'black', outlier.color = NA) + geom_jitter(width = 0.25 ,color = temp$color, size = 0.5)+
    theme(legend.position = "none", axis.text.x = element_text(angle = 30, hjust = 1)) +
    ylim(1,max_y) + xlab("Circuit") + ylab("Position") 
  b = ggplotly(b, tooltip="text")
  b$x$data[[1]]$marker$opacity = 0 
  b
})
```

### WDC Point Per Season
```{r}
renderPlotly({
colnames(driv_st)[5] = 'driverId'
res_subset <- res %>% select(driverId, constructorId, round, season)
temp <- driv_st %>% full_join(res_subset, by=c("driverId", "round", "season"))
temp <- temp %>% full_join(sche_subset, by=c("season","round"))
temp <- temp %>% full_join(cir_subset, by=c("circuitId"))
temp <- temp %>% filter(driverId == input$driver)
temp <- temp %>% mutate(text = paste0("Season: ",season,"<br>Round: ",round,"<br>Constructor: ",constructorId,"<br>Circuit: ",circuitName,", ",Location.country))
i = ggplot(temp, aes(x=round, y = points, group = season, color = as.factor(season), text = text)) + geom_line() + 
  theme_bw() + labs(color = "Year", x = "Round", y = "Points")  + theme(legend.key.size = unit(0, "cm"))
ggplotly(i, tooltip = "text")
})
```

